---
title: "Examples using the psrccensus library"
author: "suzanne"
date: "4/20/2022"
output: html_document
---
## Examples using psrccensus

The data science team asked for examples of summaries they would like to do with the psrccensus library.
The documentation is here: https://psrc.github.io/psrccensus/

•	Decennial Census: Total population by race and Hispanic/Latinx origin, 4 counties + region
o	Non-Hispanic/Latinx
	White only
	Black only
	…
o	Hispanic/Latinx
	White only
	Black only
	…
o	Also, people of color (POC)

There are a few things that are nice about scripting this. It's really easy to replicate. It automatically pulls the 4-counties and aggregates to the region.

## Population by Race/Ethnicity by County from Census
```{r get_census}
library(psrccensus)

race_county<-get_decennial_recs(geography='county', 
                   table_codes= c('P005'),
                   years= c(2010)              
                   )
race_county

write.table(race_county, "clipboard", sep="\t", row.names=FALSE)
```
I recommend doing the formatting in Excel, but if you want to learn how in R, here's how. This is kind of more of an intermediate thing, so I wouldn't worry too much about it.
```{r format_census}

library(dplyr)
library(stringr)
library(tidyr)
library(sjmisc)

race_county_formatted<-race_county %>%
   mutate(County = str_replace_all(NAME, pattern = " County, Washington", replacement = "")) %>%
   mutate(Race_Ethnicity = str_replace_all(label, pattern = "Total!!", replacement = ""))%>%    
   mutate(Race_Ethnicity = str_replace_all(Race_Ethnicity, pattern = "!!", replacement = " ")) %>%
   select(County, Race_Ethnicity, value)%>%pivot_wider(names_from=Race_Ethnicity, values_from=value)%>%
   mutate(People_of_Color=Total-`Not Hispanic or Latino White alone`)%>%
  rotate_df(cn=TRUE)
race_county_formatted
```

## Including Plots

•	ACS: Units in structure by tenure, by city by county (so cross-county city values are split across counties)



```{r find_var, echo=FALSE}
library(tidycensus)

acs_var_list<- load_variables(2020, "acs5", cache = TRUE)

```
•	PUMS: Household income by race and Hispanic/Latinx and POC (householder), 4 counties + region
Median Income
o	White only, non-Hispanic/Latinx
o	Black only, non-Hispanic/Latinx
o	…
o	Hispanic/Latinx
o	POC
```{r pums_example}

library(psrccensus)
library(magrittr)
library(dplyr)
library(srvyr)

# Obtain the data
pums19_5yr <- get_psrc_pums(5, 2019, "h", c("COUNTY","HINCP","PRACE"))

# Create a POC/Non-POC variable 
pums19_5yr %<>% mutate(POCYN=factor(case_when(PRACE=="White alone" ~"Non-POC",
                                              !is.na(PRACE) ~"POC"),
                                    levels=c("POC","Non-POC")))

# Summarize; in this case by region first and append to county-level to get all totals
# -- (Otherwise would just get a single region total)
hhinc_race_rgn <- psrc_pums_median(pums19_5yr, "HINCP","PRACE")
hhinc_race     <- psrc_pums_median(pums19_5yr, "HINCP",c("COUNTY","PRACE")) %>%
                    .[COUNTY!="Region"] %>% rbind(hhinc_race_rgn)
hhinc_poc_rgn  <- psrc_pums_median(pums19_5yr, "HINCP","POCYN")
hhinc_poc      <- psrc_pums_median(pums19_5yr, "HINCP",c("COUNTY","POCYN")) %>%
                    .[COUNTY!="Region"] %>% rbind(hhinc_poc_rgn)

```
